---
title: "Data Visualization Dashboard"
#author: "Zack Larsen"
date: "May 26, 2019"
output: 
  flexdashboard::flex_dashboard:
    theme: "journal"
    social: [ "menu" ]
    navbar:
      - { title: "About Me", href: "https://zacklarsen.github.io/" , align: right}
      - { icon: "fa-github", href: "https://github.com/ZackLarsen" , align: right}
      - { icon: "fa-linkedin", href: "https://www.linkedin.com/in/larsenzachary/", align: right}
    source: embed
---

<!-- <style> -->

<!-- body { -->
<!--   padding-top:0px -->
<!-- } -->

<!-- .navbar{ -->
<!--   visibility: hidden -->
<!-- } -->

<!-- </style> -->

```{r setup, include=FALSE}
library(pacman)
library(tidyverse)
#library(hrbrthemes)

#devtools::install_github("jeromefroe/circlepackeR")
#devtools::install_github("thomasp85/patchwork")

#https://github.com/hrbrmstr/hrbrthemes
#devtools::install_github("hrbrmstr/hrbrthemes")

p_load(flexdashboard, leaflet, leaflet.extras, dplyr, ggvis, here, conflicted, data.table, jsonlite, kableExtra, glue, geosphere, DT, networkD3, htmltools, geojsonio, magrittr, treemap, highcharter, viridisLite, ggmap, scales, ggthemes, rbokeh, data.tree, ggraph, circlepackeR, viridis, patchwork, ggraph, igraph, RColorBrewer, collapsibleTree)

#plotly

conflict_prefer("filter", "dplyr")

here::here() # "/Users/zacklarsen/Zack_Master/Projects/Dataviz/R/Divvy_Flex"

chi_data <- file.path("/Users/zacklarsen/Zack_Master/Datasets/Chicago")

stations <- fromJSON("https://feeds.divvybikes.com/stations/stations.json")

crimes <- fread(file.path(chi_data,"Crimes_sample.csv")) %>% 
  na.omit()

ward_boundaries <- geojsonio::geojson_read(file.path(chi_data,"Boundaries_Wards.geojson"), what = "sp")

```


Landing Page - Crime Map
=========================================

Column 
-------------------------------------

### Chicago crimes, plotted using leaflet. This map can be zoomed in on interactively and the station names will be displayed upon hovering over the icon.

```{r, cache=TRUE}

crimes_popup <- crimes %>% 
  mutate(
    popup_text = paste(
      sep = "<br/>",
      glue("<b><a>{`Primary Type`}</a></b>"),
      Description,
      Date,
      paste(Latitude, Longitude, sep=', ')
    )
  ) %>% 
  na.omit()

my_map <- leaflet(crimes_popup) %>%
  addTiles() %>%
  addMarkers(lat = crimes_popup$Latitude,
             lng = crimes_popup$Longitude,
             label = lapply(crimes_popup$popup_text, HTML)
             )
my_map

```

***

https://rstudio.github.io/leaflet/

- Interactive panning/zooming

- Notice the icons, when clicked, display the crime description and time when it was committed/reported


Column {.tabset}
-------------------------------------
   
### Data

```{r, cache=TRUE}

crimes %>% 
  select(`Case Number`, Date, `Primary Type`, Description) %>% 
  head(n=20) %>% 
  kable() %>% 
  kable_styling()

```   


### Crimes Involving Use of Firearm
    
```{r, cache=TRUE, eval=TRUE}

crimes %>% 
  filter(grepl("GUN",Description)) %>% 
  select(`Case Number`, Date, `Primary Type`, Description) %>% 
  arrange(Date) %>% 
  head(n=20) %>% 
  kable() %>% 
  kable_styling()

```


### Arrests Made
    
```{r, cache=TRUE, eval=TRUE}

crimes %>% 
  filter(Arrest == TRUE) %>% 
  select(`Case Number`, Date, Arrest, `Primary Type`, Description) %>% 
  arrange(Date) %>% 
  head(n=20) %>% 
  kable() %>% 
  kable_styling()

```


Statistical {.storyboard data-commentary-width=200 data-navmenu="Traditional"} 
=========================================

These visualization convey information about distributions and uncertainty.

### Boxplots are used for examining the range of values that a continuous variable takes on.

```{r, cache=TRUE, eval=TRUE}

ggplot(diamonds, aes(x=diamonds$cut, y=diamonds$price, fill=diamonds$color)) +
  geom_boxplot() + 
  scale_y_continuous(labels = dollar) + 
  #guides(fill=FALSE) + 
  ggtitle("Diamond price by cut type and color") + 
  xlab("Type") + 
  ylab("Price") + 
  theme_fivethirtyeight() + 
  coord_flip()

x1 <- ggplot(data = diamonds,
             mapping = aes(x = cut, y = price)) +
  theme_economist() + 
  scale_fill_economist()

x1 + geom_boxplot(fill = "lightpink") + 
  scale_y_continuous(label = dollar) +
  coord_flip() + 
  ggtitle("Diamond price by cut type")

```

<!-- *** -->

<!-- Here we have two groupings: the "cut" of the diamond on the y-axis and the "color", which is mapped to the "fill" attribute. -->


### Scatter Plots

```{r, cache=TRUE, eval=TRUE}

p <- ggplot(data = diamonds,
            mapping = aes(x = depth, y = price))

p + geom_point() +
  annotate(geom = "rect", xmin = 70, xmax = 80,
           ymin = 15000, ymax = 20000, fill = "red", alpha = 0.2) + 
  annotate(geom = "text", x = 70, y = 14000,
           label = "A surprisingly high \n recovery rate.", hjust = 0)

```

```{r, cache=TRUE, eval=TRUE}

figure() %>%
  ly_points(Sepal.Length, Sepal.Width, data = iris, color = Species, hover = list(Sepal.Length, Sepal.Width))

```


### Density plots are helpful in showing where the largest mass of the distribution of a continuous variable lies. They give more context than a traditional 5-figure boxplot.

```{r, cache=TRUE, eval=TRUE}

# Proportionally scaled density map
d2 <- ggplot(data = diamonds,
             mapping = aes(x = price, fill = cut, color = cut))

d2 + geom_density(alpha = 0.3, mapping = aes(y = ..scaled..))

```


### Rbokeh density

```{r cache=TRUE, eval=TRUE}

figure(width = 600, height = 400, legend_location = "top_right") %>%
  ly_hist(price, data = diamonds, breaks = 40, freq = FALSE, hover = (diamonds$price)) %>%
  ly_density(price, data = diamonds, color = "red")

```


Categorical {.storyboard data-navmenu="Traditional"} 
=========================================

These visualizations correspond to count data obtained by aggregating by certain levels or factors and count the number of observations for each category or subcategory.

### Treemap

```{r Treemap, cache=TRUE, eval=TRUE, warning=FALSE, message=FALSE}

thm <- 
  hc_theme(
    colors = c("#1a6ecc", "#434348", "#90ed7d"),
    chart = list(
      backgroundColor = "transparent",
      style = list(fontFamily = "Source Sans Pro")
    ),
    xAxis = list(
      gridLineWidth = 1
    )
  )

treemap_data <- crimes %>% 
  select(`Case Number`, `Primary Type`, Description) %>% 
  group_by(`Primary Type`, Description) %>% 
  summarise(Count = n_distinct(`Case Number`))

tm <- treemap(treemap_data, index = c("Primary Type", "Description"),
              vSize = "Count", vColor = "Count",
              type = "value", palette = rev(viridis(6)))

highchart() %>% 
  hc_add_series_treemap(tm, allowDrillToNode = TRUE,
                        layoutAlgorithm = "squarified") %>% 
  hc_add_theme(thm)

```


### Packed Bubble Charts

```{r Bubble, cache=TRUE, error=FALSE, warning=FALSE, eval=TRUE}

# https://www.r-graph-gallery.com/338-interactive-circle-packing-with-circlepacker/

# Let's use the 'flare dataset' (stored in the ggraph library)
data_edge=flare$edges
data_edge$from=gsub(".*\\.","",data_edge$from)
data_edge$to=gsub(".*\\.","",data_edge$to)
#head(data_edge)   # This is an edge list
 
# We need to convert it to a nested data frame. the data.tree library is our best friend for that:
data_tree <- FromDataFrameNetwork(data_edge)
data_nested <- ToDataFrameTree(data_tree, 
                            level1 = function(x) x$path[2],
                            level2 = function(x) x$path[3],
                            level3 = function(x) x$path[4],
                            level4 = function(x) x$path[5])[-1,-1]
data_nested <- na.omit(data_nested)
 
# Now we can plot it as seen before!
data_nested$pathString <- paste("roots", data_nested$level1, data_nested$level2, data_nested$level3, data_nested$level4, sep = "/")
data_nested$value=1
data_Node <- as.Node(data_nested)

circlepackeR(data_Node, size = "value")

```


### Collapsible Tree Diagram

```{r tree, cache=TRUE, eval=TRUE}

# https://adeelk93.github.io/collapsibleTree/

org <- data.frame(
    Manager = c(
        NA, "Ana", "Ana", "Bill", "Bill", "Bill", "Claudette", "Claudette", "Danny",
        "Fred", "Fred", "Grace", "Larry", "Larry", "Nicholas", "Nicholas"
    ),
    Employee = c(
        "Ana", "Bill", "Larry", "Claudette", "Danny", "Erika", "Fred", "Grace",
        "Henri", "Ida", "Joaquin", "Kate", "Mindy", "Nicholas", "Odette", "Peter"
    ),
    Title = c(
        "President", "VP Operations", "VP Finance", "Director", "Director", "Scientist",
        "Manager", "Manager", "Jr Scientist", "Operator", "Operator", "Associate",
        "Analyst", "Director", "Accountant", "Accountant"
    )
)

org$Color <- org$Title
levels(org$Color) <- colorspace::rainbow_hcl(11)
collapsibleTreeNetwork(
  org,
  attribute = "Title",
  fill = "Color",
  nodeSize = "leafCount",
  collapsed = FALSE
)

```


### Categorical Heatmaps

```{r, cache=TRUE, eval=FALSE}



```


Geospatial {.storyboard}
=========================================

This board is for displaying information about coordinate points or geospatial features such as countries, states, counties, rivers, elevation, etc.


### Heatmap

```{r, cache=TRUE, eval=TRUE}

leaflet(crimes) %>%
  addTiles() %>% 
  addHeatmap(lng=~Longitude,
             lat=~Latitude,
             radius = 8)

```


### Crime Totals By Ward

```{r, cache=TRUE, eval=TRUE}

ward_totals <- crimes %>% 
  select(Ward, `Primary Type`) %>% 
  group_by(Ward) %>% 
  summarise(n()) %<>% 
  mutate(count = `n()`)

ward_totals <- setNames(ward_totals$count, as.character(ward_totals$Ward))

ward_boundaries$crime_total <- ward_totals

labels <- sprintf(
  "<strong>Ward #%s</strong><br/> %g crimes committed in this ward",
  ward_boundaries$ward, ward_boundaries$crime_total
) %>% lapply(htmltools::HTML)


qpal <- colorQuantile("Reds", ward_boundaries$crime_total, n = 10)

leaflet(ward_boundaries) %>%
  addTiles() %>%
  addPolygons(fillColor = ~qpal(crime_total),
              weight = 2,
              opacity = 1,
              color = "white",
              dashArray = "3",
              fillOpacity = 0.9,
              highlight = highlightOptions(
                weight = 3, # This is the width of the dashed line
                color = "#666",
                dashArray = "",
                fillOpacity = 0.7,
                bringToFront = TRUE),
              label = ~labels,
              labelOptions = labelOptions(
                style = list("font-weight" = "normal", padding = "3px 8px"),
                textsize = "15px",
                direction = "auto")) %>% 
  addLegend(pal = qpal, 
            values = ~crime_total,
            opacity = 0.7, 
            title = "Crime Count Quantile",
            position = "topright")

```


### Clusters

```{r, cache=TRUE, eval=TRUE}

crimes_full <- crimes %>% na.omit()

leaflet(crimes_full) %>%
  addTiles() %>%
  addMarkers(
    lat = crimes_full$Latitude, 
    lng = crimes_full$Longitude,
    clusterOptions = markerClusterOptions()
  )

```


### Circle and Tile Layers

```{r, cache=TRUE, eval=TRUE}

BATTERY <- crimes[crimes$`Primary Type` == 'BATTERY',]
ROBBERY <- crimes[crimes$`Primary Type` == 'ROBBERY',]
HOMICIDE <- crimes[crimes$`Primary Type` == 'HOMICIDE',]
ASSAULT <- crimes[crimes$`Primary Type` == 'ASSAULT',]
PROSTITUTION <- crimes[crimes$`Primary Type` == 'PROSTITUTION',]
GTA <- crimes[crimes$`Primary Type` == 'MOTOR VEHICLE THEFT',]

leaflet(crimes) %>%
  # Base groups
  addTiles(group = "Default") %>%
  addProviderTiles(providers$Stamen.Toner, group = "Toner") %>%
  addProviderTiles(providers$Stamen.Watercolor, group = "Watercolor") %>% 
  addProviderTiles(providers$NASAGIBS.ViirsEarthAtNight2012, group = "NASA") %>%
  # Overlay groups
  addCircles(~BATTERY$Longitude, ~BATTERY$Latitude, group = "BATTERY") %>%
  addCircles(~ROBBERY$Longitude, ~ROBBERY$Latitude, group = "ROBBERY") %>%
  addCircles(~HOMICIDE$Longitude, ~HOMICIDE$Latitude, group = "HOMICIDE") %>%
  addCircles(~ASSAULT$Longitude, ~ASSAULT$Latitude, group = "ASSAULT") %>%
  addCircles(~PROSTITUTION$Longitude, ~PROSTITUTION$Latitude, group = "PROSTITUTION") %>%
  addCircles(~GTA$Longitude, ~GTA$Latitude, group = "GTA") %>%
  # Layers control
  addLayersControl(
    baseGroups = c("Default", "Toner", "Watercolor", "NASA"),
    overlayGroups = c("BATTERY","ROBBERY","HOMICIDE","ASSAULT","PROSTITUTION","GTA"),
    options = layersControlOptions(collapsed = FALSE)
  )

```


### Custom icon colors

```{r, warning=FALSE, error=FALSE, eval=TRUE}

#https://rstudio.github.io/leaflet/markers.html

getColor <- function(df, var) {
  quovar <- enquo(var)
  dat <- df %>% select(!!quovar)
  names(dat) <- 'x'
  lapply(dat$x, function(x) {
    if(x == 'ROBBERY') {"green"}
    else if(x == 'ASSAULT') {"orange"}
    else if(x == 'THEFT') {"blue"}
    else {"red"}
  })
}

icons <- awesomeIcons(
  icon = 'ios-close',
  iconColor = 'black',
  library = 'ion',
  markerColor = getColor(crimes, `Primary Type`)
)

leaflet(crimes) %>%
  addTiles() %>%
  addAwesomeMarkers(~Longitude, ~Latitude, icon=icons, label=~as.character(`Primary Type`))



# Alternative implementation:

# getColor <- function(df) {
#   # Need to use lapply() here instead os sapply()
#   # like the example from leaflet for r. The reason
#   # is that jsonlite is expecting a named list, not a
#   # named vector
#   lapply(df$`Primary Type`, function(`Primary Type`) {
#     if(`Primary Type` == 'ROBBERY') {"green"}
#     else if(`Primary Type` == 'ASSAULT') {"orange"}
#     else if(`Primary Type` == 'THEFT') {"blue"}
#     else {"red"}
#   })
# }
# 
# icons <- awesomeIcons(
#   icon = 'ios-close',
#   iconColor = 'black',
#   library = 'ion',
#   markerColor = getColor(crimes)
# )
# 
# leaflet(crimes) %>%
#   addTiles() %>%
#   addAwesomeMarkers(~Longitude, ~Latitude, icon=icons, label=~as.character(`Primary Type`))

```


Custom {.storyboard}
=========================================

This board is for visualizations that might not appear frequently in many examples. Some of them may be easily created by adding a few options to more popular plots, whereas some others require creative thinking and harnessing the power of visualization packages that goes well beyond pre-defined functions.

### Dumbbell (Love) plot

```{r, cache=TRUE, eval=TRUE}

#https://www.r-bloggers.com/make-multi-point-dumbbell-plots-in-ggplot2/

tibble(
  val1 = c(3, 2, 4),
  val2 = c(1, 4, 5),
  val3 = c(5, 8, 6),
  cat = factor(month.name[1:3], levels = rev(month.name[1:3]))
) -> xdf


ggplot() +
  # reshape the data frame & get min value so you can draw an eye-tracking line (this is one geom)
  geom_segment(
    data = gather(xdf, measure, val, -cat) %>% 
      group_by(cat) %>% 
      top_n(-1) %>% 
      slice(1) %>%
      ungroup(),
    aes(x = 0, xend = val, y = cat, yend = cat),
    linetype = "dotted", size = 0.5, color = "gray80"
  ) +
  # reshape the data frame & get min/max category values so you can draw the segment (this is another geom)
  geom_segment(
    data = gather(xdf, measure, val, -cat) %>% 
      group_by(cat) %>% 
      summarise(start = range(val)[1], end = range(val)[2]) %>% 
      ungroup(),
    aes(x = start, xend = end, y = cat, yend = cat),
    color = "gray80", size = 2
  ) +
  # reshape the data frame & plot the points
  geom_point(
    data = gather(xdf, measure, value, -cat),
    aes(value, cat, color = measure), 
    size = 4
  ) +
  # i just extended the scale a bit + put axis on top; choose aesthetics that work 
  # for you
  #scale_x_comma(position = "top", limits = c(0, 10)) +
  #scale_color_ipsum(name = "A real legend title") +
  labs(
    x = "Description of the value", y = NULL, 
    title = "A good plot title"
  ) +
  #theme_ipsum_rc(grid = "X") +
  theme(legend.position = "top")

```


### Hierarchical Edge Bundling

```{r}

# https://www.data-to-viz.com/graph/edge_bundling.html

# The flare dataset is provided in ggraph
edges <- flare$edges
vertices <- flare$vertices %>% arrange(name) %>% mutate(name=factor(name, name))
connections <- flare$imports

# Preparation to draw labels properly:
vertices$id=NA
myleaves=which(is.na( match(vertices$name, edges$from) ))
nleaves=length(myleaves)
vertices$id[ myleaves ] = seq(1:nleaves)
vertices$angle= 90 - 360 * vertices$id / nleaves
vertices$hjust<-ifelse( vertices$angle < -90, 1, 0)
vertices$angle<-ifelse(vertices$angle < -90, vertices$angle+180, vertices$angle)

# Build a network object from this dataset:
mygraph <- graph_from_data_frame(edges, vertices = vertices)

# The connection object must refer to the ids of the leaves:
from = match( connections$from, vertices$name)
to = match( connections$to, vertices$name)

# Make the plot
ggraph(mygraph, layout = 'dendrogram', circular = TRUE) + 
    geom_conn_bundle(data = get_con(from = from, to = to), alpha = 0.1, colour="#69b3a2") + 
    geom_node_text(aes(x = x*1.01, y=y*1.01, filter = leaf, label=shortName, angle = angle, hjust=hjust), size=1.5, alpha=1) +
    coord_fixed() +
    theme_void() +
    theme(
      legend.position="none",
      plot.margin=unit(c(0,0,0,0),"cm"),
    ) +
    expand_limits(x = c(-1.2, 1.2), y = c(-1.2, 1.2))

```


### HEB 2

```{r, cache=TRUE, warning=FALSE, eval=TRUE}

# https://www.r-graph-gallery.com/311-add-labels-to-hierarchical-edge-bundling/

# create a data frame giving the hierarchical structure of your individuals
d1=data.frame(from="origin", to=paste("group", seq(1,10), sep=""))
d2=data.frame(from=rep(d1$to, each=10), to=paste("subgroup", seq(1,100), sep="_"))
edges=rbind(d1, d2)

# create a dataframe with connection between leaves (individuals)
all_leaves=paste("subgroup", seq(1,100), sep="_")
connect=rbind( data.frame( from=sample(all_leaves, 100, replace=T) , to=sample(all_leaves, 100, replace=T)), data.frame( from=sample(head(all_leaves), 30, replace=T) , to=sample( tail(all_leaves), 30, replace=T)), data.frame( from=sample(all_leaves[25:30], 30, replace=T) , to=sample( all_leaves[55:60], 30, replace=T)), data.frame( from=sample(all_leaves[75:80], 30, replace=T) , to=sample( all_leaves[55:60], 30, replace=T)) )
connect$value=runif(nrow(connect))

# create a vertices data.frame. One line per object of our hierarchy
vertices = data.frame(
  name = unique(c(as.character(edges$from), as.character(edges$to))) , 
  value = runif(111)
) 
# Let's add a column with the group of each name. It will be useful later to color points
vertices$group = edges$from[ match( vertices$name, edges$to ) ]

#Let's add information concerning the label we are going to add: angle, horizontal adjustement and potential flip
#calculate the ANGLE of the labels
vertices$id=NA
myleaves=which(is.na(match(vertices$name, edges$from)))
nleaves=length(myleaves)
vertices$id[ myleaves ] = seq(1:nleaves)
vertices$angle = 90 - 360 * vertices$id / nleaves

# calculate the alignment of labels: right or left
# If I am on the left part of the plot, my labels have currently an angle < -90
vertices$hjust<-ifelse(vertices$angle <- 90, 1, 0)

# flip angle BY to make them readable
vertices$angle<-ifelse(vertices$angle <- 90, vertices$angle + 180, vertices$angle)


# Create a graph object
mygraph <- graph_from_data_frame(edges, vertices=vertices)

# The connection object must refer to the ids of the leaves:
from = match(connect$from, vertices$name)
to = match(connect$to, vertices$name)



ggraph(mygraph, layout = 'dendrogram', circular = TRUE) + 
  geom_conn_bundle(data = get_con(from = from, to = to), alpha=0.2, width=0.9, aes(colour=..index..)) +
  scale_edge_colour_distiller(palette = "RdPu") +
  
  geom_node_text(aes(x = x*1.15, y=y*1.15, filter = leaf, label=name, angle = angle, hjust=hjust, colour=group), size=2, alpha=1) +
  
  geom_node_point(aes(filter = leaf, x = x*1.07, y=y*1.07, colour=group, size=value, alpha=0.2)) +
  scale_colour_manual(values= rep( brewer.pal(9,"Paired") , 30)) +
  scale_size_continuous( range = c(0.1,10) ) +
  
  theme_void() +
  theme(
    legend.position="none",
    plot.margin=unit(c(0,0,0,0),"cm"),
  ) +
  expand_limits(x = c(-1.3, 1.3), y = c(-1.3, 1.3))

```
