---
title: "Divvy_Dashboard"
author: "Zack Larsen"
date: "8/25/2019"
output: 
  flexdashboard::flex_dashboard:
    theme: "journal"
    social: [ "menu" ]
    navbar:
      - { title: "About Me", href: "https://zacklarsen.github.io/" , align: right}
      - { icon: "fa-github", href: "https://github.com/ZackLarsen" , align: right}
      - { icon: "fa-linkedin", href: "https://www.linkedin.com/in/larsenzachary/", align: right}
    source: embed
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(pacman)
library(plotly)
library(magrittr)
library(tibble)

p_load(leaflet, leaflet.extras, flexdashboard, ggplot2, 
       data.table, glue, DT, kableExtra, ggvis, here, 
       conflicted, data.table, jsonlite, geojsonio, geosphere, 
       networkD3, htmltools, treemap, highcharter, viridisLite,
       ggmap, ggthemes, rbokeh, data.tree, ggraph, 
       circlepackeR, viridis, patchwork, ggraph, RColorBrewer,
       collapsibleTree, scales, janitor)

conflict_prefer("here", "here")

trips <- fread(here('data','trips.csv'), nrows = 1000)

stations <- fread(here('data','stations.csv'))

trips_with_stations <- merge(trips, stations, by.x = 'from_station_id', by.y = 'station_id')


# Data below created in ETL.R script:
by_hour <- fread(here('data','by_hour.csv'))
by_twelve_hour <- fread(here('data', 'by_twelve_hour.csv'))



# ward_boundaries <- geojsonio::geojson_read(file.path(chi_data,"Boundaries_Wards.geojson"), what = "sp")


```


Landing Page - Divvy Stations
=========================================

Column 
-------------------------------------

### Divvy stations, plotted using leaflet. This map can be zoomed in on interactively and the station names will be displayed upon hovering over the icon.


```{r landing_map, cache=TRUE, eval=TRUE, echo=FALSE, warning=FALSE}

bike_icon <- makeIcon(
  iconUrl = "/Users/zacklarsen/Zack_Master/Projects/Dataviz/R/divvy/img/Divvy-icon.png",
  iconWidth = 35, iconHeight = 35,
  iconAnchorX = 0, iconAnchorY = 10
)

station_labels <- stations %>% 
  mutate(
    popup_text = paste(
      sep = "<br/>",
      glue("<b><a>{name}</a></b>"),
      glue('Capacity: ',"<b><a>{capacity}</a></b>"),
      paste(lat, lon, sep=', ')
    )
  )

my_map <- leaflet(station_labels) %>%
  addTiles() %>%
  addMarkers(
    lat = station_labels$lat,
    lng = station_labels$lon,
    label = lapply(station_labels$popup_text, HTML),
    icon = bike_icon
    )

my_map

```

Column {.tabset}
-------------------------------------

These tabs are populated with data from Divvy's live feed, available at https://gbfs.divvybikes.com/gbfs/gbfs.json.

### Station Status


```{r status, echo=FALSE, cache=TRUE}

station_status <- fromJSON("https://gbfs.divvybikes.com/gbfs/en/station_status.json")

station_status_data <- station_status$data

station_status_data %<>% 
  as.data.frame() %>% 
  clean_names(., "small_camel")
  
colnames(station_status_data) = gsub("stations", "", colnames(station_status_data))

station_status_data %>% 
  select(StationId, NumBikesAvailable) %>% 
  arrange(-NumBikesAvailable) %>% 
  head(n=20) %>% 
  kable() %>% 
  kable_styling()

```


### Station Information

```{r info, echo=FALSE, cache=TRUE}

station_information <- fromJSON("https://gbfs.divvybikes.com/gbfs/en/station_information.json")

station_information_data <- station_information$data$stations

station_information_data %<>% 
  as.data.frame() %>% 
  clean_names(., "small_camel")
  
colnames(station_information_data) = gsub("stations", "", colnames(station_information_data))

station_information_data %>% 
  select(stationId, name, lat, lon, capacity) %>% 
  arrange(-capacity) %>% 
  head(n=20) %>% 
  kable() %>% 
  kable_styling()

```




Trips
=========================================

Column {.tabset}
-------------------------------------


### Trips

```{r, cache=TRUE, eval=TRUE, echo=FALSE}

trips %>% 
  select(start_time, end_time, bikeid, tripduration) %>% 
  head(n=20) %>% 
  kable() %>% 
  kable_styling()

```   


### Trips greater than 10 minutes
    
```{r, cache=TRUE, eval=TRUE, echo=FALSE}

trips %>% 
  select(trip_id, start_time, end_time, tripduration) %>% 
  filter(tripduration >= 10 * 60) %>% 
  mutate(duration_in_minutes = tripduration / 60) %>% 
  arrange(desc(duration_in_minutes)) %>% 
  head(n=20) %>% 
  kable() %>% 
  kable_styling()

```


### Trips on Michigan Avenue
    
```{r, cache=TRUE, eval=TRUE, echo=FALSE}

stations %>% 
  filter(grepl("Michigan", name)) %>% 
  arrange() %>% 
  head(n=20) %>% 
  kable() %>% 
  kable_styling()

```


### Stations in Evanston
    
```{r, cache=TRUE, eval=TRUE, echo=FALSE}

stations %>% 
  filter(region == 'Evanston') %>% 
  arrange() %>% 
  head(n=20) %>% 
  kable() %>% 
  kable_styling()

```


### Session Info

```{r sess}

sessionInfo()

```

Traditional {.storyboard data-commentary-width=200 data-navmenu="Traditional"} 
=========================================

These visualizations convey information about the density of certain data points geographically.


### Heatmaps

```{r, cache=TRUE, eval=TRUE}

leaflet(trips_with_stations, options = leafletOptions(minZoom = 10, maxZoom = 14)) %>%
  setView(lng = -87.71335, lat = 41.85627	, zoom = 10) %>% 
  addTiles() %>% 
  addHeatmap(lng=~lon,
             lat=~lat,
             radius = 8)

```


Histograms {.storyboard data-commentary-width=200 data-navmenu="Other"} 
=========================================

These visualizations convey information about trips and riders.


### Histogram by hour


```{r histo_hour, cache=TRUE, warning=FALSE, message=FALSE}

by_hour_data <- by_hour %>% 
  mutate(start_hour = as.factor(start_hour))

ggplot(by_hour_data, aes(x = start_hour, y = n)) +
  geom_bar(fill = "dodgerblue", width = 0.6, stat = "identity") +
  facet_wrap(~dow, ncol = 2) +
  scale_x_discrete(breaks = seq(0,24,2)) +
  theme_minimal() +
  scale_x_discrete() + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

```


### Histogram by AM/PM


```{r histo_am_pm}

by_twelve_hour_data <- by_twelve_hour %>% 
  mutate(twelve_hour2 = as.factor(twelve_hour2))

ggplot(by_twelve_hour_data, aes(x = twelve_hour2, y = n)) +
  geom_bar(fill = "dodgerblue", width = 0.6, stat="identity") +
  facet_wrap(~dow, ncol = 2) +
  scale_x_discrete(breaks = seq(0,24,2)) +
  theme_minimal() +
  scale_x_discrete() + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

```
